--- 
title: "Waiting"
author: "Steven Rimmer"
date: "16 April 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Introduction

Code to demonstrate the link between the waiting times in a Poisson process to
the Gamma and Exponential distributions.

* Simulate an approximation to a Poisson process using discrete time steps
* Measure the waiting time to each kth event
* Group the output to plot distribution 
* Compare the grouped output to waiting times drawn from: 
* The Gamma distribution 
* The Exponential distribution in the special case where k = 1, the waiting time between each Poisson events

## Parameters and initialisation

Clear the memory of objects stored from previous work and call libraries: 
```{r initialisation}
rm(list=ls())	# Remove objects currently in the workspace 
require(stats)							# Call the stats package for gamma and exponential distributions 
library(tibble)
library(ggplot2)
library(dplyr) 
```

Set the timer: 
```{r tic} 
tic = Sys.time() 
```

Initiaise parameters:
```{r parameters} 
p  = 0.001                        # Ratefor poisson process 
n  = 10e6                         # Number of time steps 
k = 5                            # Waiting time for the kth arrival simulated 
k_count = 0                       # Counter to keep count up to kth-event 
event <- array(dim=c(n))					# Initialise an n-array to record when events occur 
count_event <- array(dim=c(n))					# Initialise an n-array to record the cumulative count of kth-events 
clock_event <- array(dim=c(n))					# Initialise an n-array to record the time at which each kth-event happened 
plot_bins <- n*p/k/10 
```

## Simulation

Simulate events occurring for each time step using uniform distribution: 
```{r simulate} 
event = runif(n,0,1) < p 
```

We want to Consider the first time step separately.

```{r step_one} 
if (event[1] == 0) {						# If the first time step does not have an event: 
  count_event[1] = 0 } else {				#	Set the cumulative count to zero 
                                        # If the first time step does contain an event:
  if (k_count == (k-1)) {					# 	And if we have reached the kth event: 
    count_event[1] = 1					#		Set the cumulative kth-count to one
    clock_event[1] = 1					#		Record the first kth-event happening at t = 1 
    } else {						#	Else if we have not reached the kth event: 
count_event[1] = 0					#		Set the cumulative count to zero 
k_count = k_count +1}					#		Increase the count of events seen before reaching the kth 
    } 
for (i in 2:n) {						# Set up a loop to consider the subsequent time steps in turn 
  if (event[i] == 0) {					# If this time step does not have an event: 
  count_event[i] = count_event[i-1]			# 	Increase the cumulative count by one 
    } else {						# If this time step does contain an event: 
      if (k_count == (k-1)) {					#	And if we have reached the kth event: 
        count_event[i] = count_event[i-1] + 1			#		Increase the cumulative count by one 
        clock_event[count_event[i]] = i				#		Record the time step at which this event occurred 
        k_count = 0						#		Reset the k-counter to start counting again
        } else {						#	Else if we have not reached the kth event: 
          k_count = k_count + 1					#		Increease the k-counter by one 
          count_event[i] = count_event[i-1]			#		Maintain the cumulative event count 
          }} }
tot_event = count_event[n]					# Record the total number of kth-events 
wait_time <- array(dim=c(tot_event))				# Inialise an array to record the wait time to each kth-event 
wait_time[1] = clock_event[1]-1					# Record the wait time to the first kth-event 
for (i in 2:tot_event) { 					# Set-up a loop to cycle through each kth-event
  wait_time[i] = clock_event[i]-clock_event[i-1] }	# Record the waiting for the kth-event

toc = Sys.time()						# End timer 

waiting <- tibble(event_num =1:tot_event,
                  clock_event = clock_event[1:tot_event],
                  wait_time=wait_time[1:tot_event]) 
rm(clock_event,count_event,event)
```

```{r plot} 
(histplot <-
ggplot(waiting)+geom_histogram(aes(wait_time),breaks=seq(0,10*n*p/k,by=plot_bins)))
```

``` {r comparison} 
wait_bucket <- array(dim=c(tot_event))				# Initialise array to bucket simulated waiting times for graphical output 
wait_bucket = ceiling(wait_time/plot_bins)				# Allocate simulated wait times to a bucket 
num_bucket = max(wait_bucket)
count_bucket = array(dim=c(num_bucket)) 
for (i in 1:num_bucket) {count_bucket[i]=sum(wait_bucket==i)}  
#ggplot_build(histplot)$data[[1]] 
#count_bucket 
```

``` {r compare} 
compare_plot <- tibble(centre = seq(0.5,(length(count_bucket)-0.5))*plot_bins, 
                       count = count_bucket,
                       gamma = (n*p*plot_bins/k)*dgamma(centre,shape=k,rate=p), 
                       expdist = (n*p*plot_bins)*dexp(centre,rate=p)) 
ggplot(compare_plot)+
  geom_col(aes(centre,count),width=plot_bins)+ 
  geom_line(aes(centre,gamma),colour="red",size=1.2)+
  if (k==1) {geom_line(aes(centre,expdist),colour="blue",size=1.2)} 
```
